---
title: "Activity 1"
author: "obe432"
date: "2024-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(tidyverse)
```


```{r load dataset}
View(phc_data)
```


1.number of obsevartion and number of variables
```{r number of rows and columns}
ncol(phc_data)
nrow(phc_data)
```

2. identify all variables with missing data
```{r missing value (N/A)}
colnames(phc_data)

colSums(is.na(phc_data))

```

3. type of variable
```{r structure of the dataset}
glimpse(phc_data)  #or
sapply(phc_data, class) 
```



Question 4, we have to select the column or variabke we will be using for the assignment. We have to change some label to factors or character for more understanding.

```{r selecting the type of variable and mutating some of the variable type to labels for understanding}
phc_data_1 <- phc_data %>% 
  select(nqid, region, a11d, p02, p17, p10) %>% 
  mutate(across(c(region, a11d, p17, p10), sjlabelled::as_label)) 

phc_data_1
```

types of variable and missing values in the new data
```{r}
sapply(phc_data_1, class)

colSums(is.na(phc_data_1))

```


4. without applying weight to the data generate 

*number of male and female by region
```{r number of a11d by region }
gender_region_count <- phc_data_1 %>%
group_by(region, a11d) %>%
summarise(count = n()) %>%
arrange(region)
gender_region_count

```
*number of male and female by sector of employment
```{r no of a12d by p17}
gender_sector_count <- phc_data_1 %>%
group_by(p17, a11d) %>%
summarise(count = n()) %>%
arrange(p17)
gender_sector_count
```



*number of girls below 12 years who are married

first group them into a categorical data 
```{r catogerical data ( informal marriage, married, never married, married before)}
 
phc_data_2 <- phc_data_1 %>%
  mutate(marital_status = case_when(
    p10 == "Informal/living together" ~ "Informal marriage",
    p10 %in% c("Married (Civil/Ordinance)", "Married (Customary/Traditional)", "Married (Islamic)", "Married (Other type)") ~ "Married",
    p10 == "Never married" ~ "Never married",
    p10 %in% c("Separated", "Divorced", "Widowed") ~ "Married before"
  )) %>%
   mutate(marital_status = factor(marital_status, levels = c("Informal marriage", "Married", "Never married","Married before")))
  
phc_data_2

```


```{r no of female < 12yrs who are married}
girls_married_count <- phc_data_2 %>%
  dplyr::filter(a11d == "Female", p02 < 12, marital_status == "Married") %>%
  summarise(count = n())

girls_married_count

```


